Performance portability of an intermediate-complexity atmospheric research model in coarray Fortran

### 2017-08-16 body.tex
Revise intro & add omnivore's dilemma citation

\section{Introduction}
\subsection{Motivation and Background}
Since the Fortran 2008 standard was published in 2010~\cite{iso2010fortran}, a handful of studies in the open literature
have provided encouraging assessments of Fortran's intrinsic parallel programming model, \gls{caf}, running in research
applications running at scale~\cite{preissl2011multithreaded,garain2015comparing,mozdzynski2015partitioned}.   One of the
most attractive features of \gls{caf} is the ability to express parallel algorithms once within the confines of a single
language standard without embedding the use of compiler directives or communication layers defined outside the Fortran
Fortran language standard (e.g., no \gls{mpi} or OpenMP.  In theory, this should leave the choice of communication layers
as a link-time decision.  Ideally, an application developer writes a standard Fortran-only application, compiles it with any
standard-conforming compiler, and possibly reserves the option to incorporate any one (or more) of several communication layers
at link-time.

Such flexibility poses the computing equivalent of what food author Michael Pollen termed ``The Omnivore's Dilemma:'' if I
belong to a species that owes some of its evolutionary advantage to being able to eat a wide variety of foods, what food is
best for me to eat?  Similarly, if I can express my parallel algorithm once using \gls{caf} and then consume cycles atop a
multitude of software stacks and hardware platforms, where best to consume cycles and using which supporting software stack?
Here we report the results of an initial study of the current options for compiling, linking, and executing a \gls{mini-app}
designed to be representative of the parallel numerical algorithms and physics employed in the \gls{icar} being developed at
the \gls{ncar}.

\gls{icar} simulates the motion of the atmosphere at kilometer length scales and produces flow patterns with a fidelity that is
attractive to the hydrology community, which studies surface water.  The animation in Figure~\ref{figure:icar} depicts atmospheric processes over North America.

>>>>

\section{Introduction}
\subsection{Motivation and Background}
Since the Fortran 2008 standard was published in 2010~\cite{iso2010information}, a handful of studies in the open literature
have provided encouraging assessments of Fortran's intrinsic parallel programming model, \gls{caf}, in research
applications running at scale~\cite{preissl2011multithreaded,garain2015comparing,mozdzynski2015partitioned}.   One of the
most attractive features of \gls{caf} is the ability to express parallel algorithms once within the confines of a single
language standard without embedding compiler directives or referencing communication layers that are not part of the
language standard (e.g., no \gls{mpi} or OpenMP).  In theory, this should leave the choice of communication layers
as a link-time decision.  Ideally, an application developer writes a standard, Fortran-only application, compiles it with any
standard-conforming compiler, and possibly reserves the option to incorporate any one (or more) of several communication layers
at link-time.

Such flexibility poses the computing equivalent of what food author Michael Pollan termed ``The Omnivore's Dilemma:'' if I
belong to a species that owes some of its evolutionary advantage to being able to eat a wide variety of foods, what food is
best for me to eat?~\cite{pollan2006omnivore}  Similarly, if I can express my parallel algorithm once using \gls{caf} and then consume cycles atop a
multitude of software stacks and hardware platforms, where best to consume cycles and using which supporting software stack?
Here we report the results of an initial study of several current options for compiling, linking, and executing a \gls{mini-app}
designed to be representative of the parallel numerical algorithms and physics employed in \gls{icar}.

\gls{icar} simulates the motion of the atmosphere at kilometer length scales and produces flow patterns with a fidelity that is
attractive to the hydrology community, which studies surface water.  The animation in Figure~\ref{figure:icar} depicts atmospheric
processes over North America as simulated by \gls{icar}.
